\documentclass[12pt]{article}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[toc,page]{appendix}
\usepackage{scrextend}
\usepackage{pgfplots}
\usepackage{listings}
\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{subcaption}
\expandafter\def\csname ver@subfig.sty\endcsname{}
\usepackage{svg}
\usepackage{amssymb}
\usepackage{courier}

\setlength\parskip{5mm}
\widowpenalty10000
\clubpenalty10000
\setlength\parindent{0pt}
\captionsetup{width=0.8\textwidth}

\lstset{ %
  language=C,
  directivestyle={\color{black}},
  emph={int,char,double,float,unsigned,struct},
  emphstyle={\color{blue}},
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  escapeinside=\`\`,               % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\lstset{basicstyle=\ttfamily,breaklines=true}

\begin{document}

\newgeometry{margin=1in}

\begin{center}

	UNIVERSITY of CALIFORNIA\\
	\vspace{2mm}
	SANTA CRUZ\\
	\vspace{10mm}
	{\large\textbf{\textsc{SOLVING BOUNDARY VALUE PROBLEMS\\EFFICIENTLY USING COMPUTERS}}}\\
	\vspace{10mm}
	A thesis submitted in partial satisfaction\\of the requirements for the degree of\\
	\vspace{5mm}
	\textsc{BACHELOR OF SCIENCE}\\
	\vspace{3mm}
	in\\
	\vspace{3mm}
	\textsc{PHYSICS}\\
	\vspace{5mm}
	by\\
	\vspace{2mm}
	Daniel A. Bittman\\
	May 29, 2016\\

	\vspace{2in}

	The thesis of Daniel A. Bittman is approved by:\\
	\vspace{1in}

	\rule{2.5in}{2pt}\hfill
	\rule{2.5in}{2pt} \\
	Professor Jason Nielsen \hfill Professor David P. Belanger\\
	Thesis Adviser \hfill Thesis Coordinator\\

	\vspace{15mm}
	\rule{2.5in}{2pt}\\
	Professor Robert Johnson\\
	Chair, Department of Physics

\end{center}

\pagenumbering{gobble}
\clearpage
\setcounter{page}{1}
\renewcommand{\thepage}{\roman{page}}

\begin{center}
	\textsc{{\Large Copyright and Software License}}\\
	\vspace{1in}
	\textsc{this document and software described herein are\\\vspace{5mm}copyright} \copyright $\;$\textsc{2016\\daniel a. bittman}\\
	\vspace{1in}
	\textsc{The software described herein, including excerpts and other code, is provided with the following license:}

	MIT License

Copyright \copyright $\;$ 2016 Daniel A. Bittman

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the ``Software''), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED ``AS IS'', WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.

\end{center}
\clearpage

\begin{addmargin}[8em]{8em}
\begin{center}

	\textbf{Dedication}
	\vspace{50mm}

	To my loving parents, Steve Bittman and Ellen Klima, for their unending support
	and unwavering belief in my ability to succeed.

	\vspace{100mm}
	Also for Frankie.
\end{center}

\begin{addmargin}[6.2em]{8em}
\begin{verbatim}
      \    /\
       )  ( ')
      (  /  )
       \(__)|
\end{verbatim}
\end{addmargin}
\end{addmargin}
\clearpage
\begin{center}
	\textbf{Acknowledgements}
	\vspace{50mm}

	A huge thank you to my thesis adviser, Jason Nielsen, for providing the opportunity
	to work on this interesting project, and for the many interesting conversations over
	the quarters past.

	\vspace{10mm}
	The Storage Systems Research Center at UC Santa Cruz for some of their machines in order
	to test the performance of this system.

\end{center}
\clearpage
\restoregeometry
\tableofcontents
\clearpage
\listoffigures
\listoftables
\clearpage
\renewcommand{\lstlistlistingname}{List of Source Code Listings}
\lstlistoflistings
\clearpage

\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\section*{Abstract}

\onehalfspacing

There are many problems in physics which involve solving a set of equations that define behavior
inside a region. One such set of problems, among many, are those of electrostatics inside
a region given some boundary conditions. Such problems, along with other
problems, are commonly found in simple and analytically solvable forms in undergraduate physics classes.
Unfortunately, these problems can quickly become intractable to solve analytically as their complexity
increases. Complex boundary conditions in strangely shaped regions can easy become impossibly difficult to
solve. The solution can instead be modeled on a computer in order to compute a numerical solution
in place of an analytical one.

I have written a software package which is C library capable of solving such a given problem, and a frontend
Python program which can parse a human-readable configuration file that describes such a boundary value problem. 
The Python code sets up the correct initial state such that the C library can solve it, after which the Python program
displays the final result as well as calculates statistics about the solving process. I have utilized effective
programming techniques designed for current generation hardware 64-bit x86 processors in order to improve
the performance of the program, in addition to implementing a multi-threaded version of the solving program
to explore the use of multi-core processors in solving these problems.

I found that the program correctly matched an analytical solution to a simple test problem, and that it correctly
predicted the $1/r^2$ potential found from an electric dipole. I was able to optimize the program so that it
efficiently found these results, and I was able to successfully leverage multi-threading to significantly increase
the performance, albeit with a degradation in simulation accuracy.

\clearpage
\doublespacing
\include{intro}

\include{proced}

\include{results}

\include{conclusion}


\clearpage
\begin{appendices}
\section{Program Usage}

Compiling the engine requires make and a modern gcc (or clang, although clang is untested), and can
be done by simply invoking \texttt{make}. This will also create an executable file, \texttt{solve}. This
is a script that invokes the python example frontend so that it may be used.

For more details on using the python frontend or the C backend engine, see the man pages (reproduced on the
following pages). There are several example files, including \texttt{example\_sin.txt} and \texttt{example\_dipole.txt}.
For more details on writing configurations for a desired simulation or experiment, see the man pages.

The source code also contains a directory named \texttt{tests}, which contains python code to generate verifier data
files for the \texttt{example\_sin} configuration. Finally, all of the data written on in this document are present
in the source directory, under \texttt{results}. The source code for the C library is in \texttt{engine}, and the
source code for the python example frontend is under \texttt{frontend}. The source code and git revision history
is available at \url{http://github.com/dbittman/statics-solver}.

	\begin{singlespacing}
\newgeometry{margin=1in}
\include{man-1}
\include{man-3}
\restoregeometry
	\end{singlespacing}
\clearpage

\include{opt}


\include{impl}

\end{appendices}

\clearpage

\section*{Software Used}

\begin{enumerate}
	\item \texttt{gnuplot}. \url{http://www.gnuplot.info}
	\item \texttt{gcc}, and associated development toolchain. \url{http://gcc.gnu.org}, \url{https://www.gnu.org/software/binutils}, etc.
	\item \texttt{python 3}. \url{https://www.python.org}, including packages:
		\begin{enumerate}
			\item \texttt{matplotlib}
			\item \texttt{numpy}
			\item \texttt{ctypes}
			\item \texttt{pickle}
		\end{enumerate}
\end{enumerate}

\clearpage
\bibliographystyle{plain}
\onehalfspacing
\bibliography{main}





\end{document}

